Shrinking Characteristics of Precision Matrix Estimators: An Illustration via Regression

### 2019-04-09 03-SCPME.Rmd
only discussion left

Figure \@ref(fig:scpmesim2) displays the model error for each method by dimension of the design matrix. Here we took the sample size equal to $n = 100$ and the response matrix dimension $r = 10$ with each data generating procedure replicated a total of 20 times. Note `OLS` and `oracleS` are not shown due to extremely poor performance.

We find that...

In high dimensional settings, `shrinkBO` (penalty: $\lambda\left\| \Omega[\hat{\Sigma}_{xy}, I_{p}] \right\|_{1}$) and `shrinkB` (penalty: $\lambda\left\| \Omega\hat{\Sigma}_{xy} \right\|_{1}$) perform increasingly well relative (and superior) to the others when $p >> n$. It's not surprising that `shrinkBO` performs the best because the embedded assumptions most closely match the data generating model.

@@

When $n > p$ we see that `shrinkB` performs terribly. Interestingly, `shrinkBO` is still one of the better performing estimators -- though worse than both `lasso` and `ridge`.

It is interesting that `oracleO` performed worse than `oracleS` in the final table. It appears that estimating the covariance matrix $\Sigma_{xy}$ well is *more* important than estimating $\Omega$ well.

@@

\vspace{0.5cm}
```{r, eval = T, echo = F, fig.cap = "ME by P (N = 100, R = 10, sparse = 0.5, sigma = tridiag)"}

# ME table (N = 100, sparse = 0.5, sigma = "tridiag", P = 150, R = 10)
sim6 %>% filter(N == 100, sparse == 0.5, sigma == "tridiag", P == 150, R == 10) %>% group_by(model) %>% summarise(mean = mean(Error), sd = sd(Error)) %>% arrange(mean) %>% pander(caption = "ME by P (N = 100, R = 10, sparse = 0.5, sigma = tridiag)")

@@

Here we compare the `shrinkB` and the `shrinkBO` estimators.

(same data generation except the parameters are different this time...)

We consider the response dimension as 10.

For each of the estimators, $\Sigma_{xy}$ is estimated by the sample covariance matrix $\hat{\Sigma}_{xy}$ times some constant factor $k \in (0.1, 0.2, ..., 0.9, 1)$. We are shrinking $\hat{\Sigma}_{xy}$ by a constant to determine if there are any benefits to doing so.

In the high dimensional settings, it does appear that shrinking the sample covariance matrix $\hat{\Sigma}_{xy}$ by a constant factor helps in reducing MSPE. This is indicated by the fact that the optimal `const` hovers around 0.5 for each of the estimators `shrinkB` and `shrinkBO`.

In the low dimension settings, it appears that shrinking the covariance matrix is not beneficial.

The type of oracle precision matrix (tridiagonal or compound symmetric) does not appear to be too influential with regards to the affect of shrinking the covariance matrix.

The larger the dimension of $Y$, the more beneficial shrinkage will likely be (not that surprising)

@@

>>>>

Figure \@ref(fig:scpmesim2) displays the model error for each method by dimension of the design matrix. Here we took the sample size equal to $n = 100$ and the response matrix dimension $r = 10$ with each data generating procedure replicated a total of 20 times. Note `OLS` and `oracleS` are not shown due to extremely poor performance. We find that in high dimensional settings, `shrinkBO` and `shrinkB` perform increasingly well relative to the others as the predictor dimension increases with `shrinkBO` performing the best. Interestingly, when $n > p$ is still one of the best-performing estimators - though worse than both `lasso` and `ridge` - but the performance of `shrinkB` decreases drastically (plot not shown).

@@

The table below shows the average model error for each estimator when $n = 100$, $p = 150$, and $r = 10$.

@@

\newpage
```{r, eval = T, echo = F, fig.cap = "ME by P (N = 100, R = 10, sparse = 0.5, sigma = tridiag)"}

# ME table (N = 100, sparse = 0.5, sigma = "tridiag", P = 150, R = 10)
sim6 %>% filter(N == 100, sparse == 0.5, sigma == "tridiag", P == 150, R == 10) %>% group_by(model) %>% summarise(error = mean(Error), sd = sd(Error)) %>% arrange(error) %>% pander(caption = "ME by P (N = 100, R = 10, sparse = 0.5, sigma = tridiag)")

@@

This final simulation was run and explored due to the fact each of the SCPME regression estimators require an estimation of the cross-covariance matrix. In the previous simulations, we used the maximum likelihood estimator. However, because the data-generating procedures construct settings that are inherently sparse, we were interested to determine if additional shrinkage of the *cross-covariance* matrix would also prove beneficial. For instance, in a number of settings, we were finding that the estimator `oracleO` ($\beta = \Omega_{x}^{*}\hat{\Sigma}_{xy}$) was performing worse than `oracleS` ($\beta = \hat{\Omega}_{x}\Sigma_{xy}^{*}$) which perhaps suggests that estimating the covariance matrix $\Sigma_{xy}$ well is *more* important than estimating $\Omega$ well. This simulation explores that theory a bit further.

The data-generating procedure for this simulation is similar to previous one but here we take $n = 100$, $p = 200$, $r = 10$, *and* we additionally multiple the cross-covariance sample estimator by a factor of $k$ where $k \in (0.1, 0.2, ..., 0.9, 1)$. Figure \@ref(fig:scpmesim7) plots the MSPE for `shrinkB` for each tuning parameter, $\lambda$, and constant, $k$, pair. Figure \@ref(fig:scpmesim9) plots the MSPE similarly for `shrinkBO`.

Interestingly, we find that in this high dimension setting, it does appear that shrinking the sample covariance matrix by a constant factor helps the overall prediction performance of the SCPME estimators. This is indicated by the fact that the optimal constant, $k$, hovers between 0.3 and 0.6 for each of the estimators. We also performed a simulation in low dimensions (not shown) but we did not see the same benefit.

@@

\newpage
